Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-10-08 paper.tex
some notes on checkpointing

            Since Spark does not natively interface with HPC schedulers, 
            the use of overlay Spark clusters, similar to HoD, has been 
            investigated for Spark applications running on HPC. The reasons 
            listed for HoD's failure include poor 
            data locality when deploying a dynamic cluster on a shared HDFS 
            instance, inability to resize cluster resulting in idle nodes and 
            high cluster allocation latency. To address the issue of high 
            cluster allocation, pilot-agent scheduling strategies can be used.
            Pilot-agent scheduling strategies allow for clusters to grow 
            dynamically and to add nodes when required. This relieves users, 
            in the case of Spark, from having to wait until all nodes become 
            available to begin execution. Both papers \note{jha and baer} use 
            this strategy. Furthermore, pilot-agent scheduling could be adapted
            to enable smart cluster resizing. Conversly, maintaining data 
            locality using a shared HDFS instance will remain an issue.

        \section{Performance enhancement strategies}\label{other}
            scientific: workflow checkpointing (looking at file hashes)

@@

>>>>

            Since Spark does not natively interface with HPC schedulers, 
            the use of overlay Spark clusters, similar to HoD, has been 
            investigated for Spark applications running on HPC. The reasons 
            listed for HoD's failure include poor 
            data locality when deploying a dynamic cluster on a shared HDFS 
            instance, inability to resize cluster resulting in idle nodes and 
            high cluster allocation latency. To address the issue of high 
            cluster allocation, pilot-agent scheduling strategies can be used.
            Pilot-agent scheduling strategies allow for clusters to grow 
            dynamically and to add nodes when required. This relieves users, 
            in the case of Spark, from having to wait until all nodes become 
            available to begin execution. Both papers \note{jha and baer} use 
            this strategy. Furthermore, pilot-agent scheduling could be adapted
            to enable smart cluster resizing. Conversly, maintaining data 
            locality using a shared HDFS instance will remain an issue. 
            \tristan{The section is interesting, but it lacks a bit of context: one would like to 
            understand why Spark and Hadoop developed their own scheduler instead of
            reusing the ones available on HPC. And what are the main characteristics of these
            schedulers: data locality should be mentioned here.}

        \section{Checkpointing}\label{other}

            Scientific workflow engines employ other strategies for improving   
            performance. As scientific worflows are compute intensive, it is    
            necessary for workflow engines to avoid repeating steps that have   
            completed successfully. This is achieved by scientific workflow     
            engines through the process of checkpointing. Workflow engines such 
            as Nipype, PSOM, AA, Pydpiper, Nextflow, Taverna, Pegasus, Galaxy,  
            Toil, Make and LONI all provide a mechanism to resume a pipeline    
            from last successful state. This can be done either by detecting    
            changes in the task definition (e.g. Nipype, PSOM, AA, Pydpiper,    
            Nextflow), checking if modification of input                        
            data is more recent than previously stored output data (e.g. Make), 
            or requesting that the user manually restart execution from         
            selected task (e.g. LONI).

            Big Data frameworks such as MapReduce, Spark and Dask do not have   
            any form of checkpointing functionality outside of a given          
            execution. For instance, it is possible to checkpoint an RDD, but   
            that RDD will only be accessible within the application's context   
            and will not be shared with future executions. As failures do       
            happen and individual neuroimaging tasks may take a non-negligable  
            amount of time to execute, a mechanism to checkpoint workflows      
            will need to be added to any BigData framework used for             
            neuroinformatics.                                                   
            \note{dont know how Taverna, Pegasus, Galaxy or Toil achieves this} 
            \note{openMole says it checks to see if outputs exist. similar to   
            make?}                                                              
            \note{what about SPM, FastR??}

@@

        \section{Evaluation of BigData Frameworks for Scientific Computing}     
            The BigData frameworks have sparked significant interest within the 
            scientific community, leading to research on the efficiency of      
            these frameworks for scientific computing. ~\cite{valduriez} found  
            that Spark scales better than expected with longer tasks. This is   
            because                                                             
            there is less I/O occurring and the scheduler is not as overloaded. 
            Additionally, ~\cite{Boubela} found that Spark in conjunction with  
            GPU acceleration greatly reduced the processing time of large       
            neuroimaging datasets.                                              
            As scientific, specifically neuroimaging workflows are typically    
            comprised of long tasks, it                                         
            is crucial that BigData frameworks scale well with tasks of longer  
            durations.

            To determine how different BigData frameworks compare in their 
            efficiency for the processing of scientific workflows, 
            ~\cite{ariel} compared Spark, Dask, SciDB, Myria and TensorFlow on
            a neuroimaging and astronomy use case. It was found that Dask, 
            Myria and Spark performed consistently better and had comparable 
            performance. Dask had poorer performance when data did not fit in
            memory. It was also found that Spark exhibited slower performance 
            when the individual images are larger due to the lack of overlap 
            between shuffle and map stages \note{revise this}. Overall it was
            found that BigData Frameworks still required some adaptations to be
            effectively used as scientific workflow engines.
